Cosmic web alignments with the shape, angular momentum and peculiar velocities of dark matter halos

### 2014-02-17 paper/main.tex
feeding the introduction

\title[Cosmic web alignments with halos]{Cosmic web alignments with the shape, angular momentum and peculiar velocities of dark matter halos}  
\author[J.E. Forero-Romero et al.]{
\parbox[t]{\textwidth}{\raggedright 
  Jaime E. Forero-Romero$^{1}$,
  Sergio Contreras$^{2}$,
  Nelson Padilla$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$Uni A
$^{2}$Uni B
}
\maketitle

@@

We study the alignment of dark matter halos with the cosmic web as described by
the tidal and velocity shear fields. We focus on the alignment of their
shape, angular momentum and peculiar velocities. We use a cosmological N-body
simulation that allows to study dark matter halos spanning almost five orders
of magnitude in mass ($10^{9}$-$10^{14}$) \hMsun over spatial scales
of $(0.5$-$1.0)$\hMpc to define the cosmic web. After varying the
numerical parameters in our experiments we find that the strongest
alignment is measured for the halo shape with respect to the web
defined by the tidal field, with a stronger signal as the halo mass
increases. For the angular momentum we find weaker alignment signals for
halos more massive than $10^{12}$\hMsun both in the tidal and velocity
shear descriptions. The peculiar velocities of halos show a strong
alignment signal for the tidal web at all halo masses, while in the
velocity shear description the signal is weaker an only present for
halos less massive than $10^{12}$\hMsun. We also find that the main
directions of stronger tidal tension and velocity shear coincide for
halos below $10^{12}$\hMpc but are perpendicular to each other for
halos more massive than this threshold.

@@

The formation of large scale structure in the Universe proceeds by
gravitational instability. This is an intrinsic anisotropic
process. [Pancake formation ...]

Numerical simulations allow the exploration of a deep nonlinear
regime, following the formation of virialized dark matter halos, which
host observable galaxies. Some characteristics of these halos have to
be linked to the environment in which they formed. One way to studty
the connection halo-environment is to quantify the degree of
correlation between different halo properties and their surrounding
density and velocity fields.

Among the properties that are of interest we find the shape, spin and
peculiar velocity. [Relevanc of dark matter halo shapes ...]
[Relevance of halo spin ...] [Relevance of peculiar velocities]

There is large tradition of theoretical precedents for the
measurements of shape and spin alignment. In this \documentname we
review most of these studyes and offer our own study with
complementary numerical techniques and simulations. [Brief summary of
  the appendix ...]

We also present the results for the alignment of peculiar velocities
with large scale structure. Such a study has received less atention in
the literature. However, we consider that it presents relevant
insights for the study of cosmic flows.

@@

\begin{figure*}
\includegraphics[width=0.8\textwidth]{Fig0.pdf}
\caption{Fraction of halos in a mass bin that show a preferential
  alignment with respect to an eigenvector. Each row presents on of
  the three properties studied in this paper: shape (major axis),
  angular momentum and  peculiar velocity. Left column presents the
  results against the Vweb   and the right the Tweb. Strong colors
  refer to $256^3$ grid resolutions and lighter colors to a $512^3$
  grid. The thick black horizontal line at $0.33$ corresponds to the
  expected fraction for a random vector field.
\label{fig:preferential}}
\end{figure*}

@@

\begin{figure}
\includegraphics[width=0.40\textwidth]{web.pdf}
\caption{Median of the interweb alignment for the two grid
  resolutions as a function of the dark matter mass corresponding to
  the halos where the measurement was made. The error bars indicate
  the lower and upper quartiles (Figure in color in the web version).
\label{fig:interweb}} 
\end{figure}

@@

Possible conclusion. Prolateness follow concentration. In the Tweb
Angular momentum alignment is influenced by spin at higher masses. In
the Vweb angular momentum alginment is influenced by all factors. In
the peculiar velocity alginment only spin seemss to have an effect.

\begin{figure*}
\includegraphics[width=0.90\textwidth,angle=90]{all_cuts.pdf}
\caption{The alignment is always taken with respect to the third eigenvector.}
\end{figure*}
\section{Discussion}
\label{sec:discussion}

>>>>

\title[Halo alignments with the cosmic web]{Cosmic web alignments with
  the shape, angular momentum and peculiar velocities of dark matter
  halos}   
\author[J.E. Forero-Romero et al.]{
\parbox[t]{\textwidth}{\raggedright 
  Jaime E. Forero-Romero$^{1}$,
  Sergio Contreras$^{2}$,
  Nelson Padilla$^{2}$ 
}
\vspace*{6pt}\\
$^{1}$Uni A
$^{2}$Uni B
}
\maketitle

@@

We study the alignment of dark matter halos with the cosmic web as described by
the tidal and velocity shear fields. We focus on the alignment of their
shape, angular momentum and peculiar velocities. We use a cosmological N-body
simulation that allows to study dark matter halos spanning almost five orders
of magnitude in mass ($10^{9}$-$10^{14}$) \hMsun over spatial scales
of $(0.5$-$1.0)$\hMpc to define the cosmic web. After varying the
numerical parameters in our experiments we find that the halo shape
and the tidal field web presents the strongest alignment with a signal
that gets stronger as the halo mass increases. For the angular momentum we
find alignment signals only halos more massive than $10^{12}$\hMsun
both in the tidal and velocity shear descriptions. The peculiar
velocities of halos show a strong alignment signal for the tidal web
at all halo masses, while in the velocity shear description the signal
is weaker an only present for halos less massive than
$10^{12}$\hMsun. We also find that the main directions of stronger
tidal tension and velocity shear coincide for halos below
$10^{12}$\hMpc but are perpendicular to each other for halos more
massive than this threshold.

@@

There is a long observational tradition stuying galactic properties
as a function of its large scale environment. In these situations the environment
definition is based on quantities easily accesible to observations
such as the number density. With the advenment of large
surveys the visibility of the cosmic web is clear. As a consequence,
in order to capture this filamentarity, the environment definition
started to be more complex, involving shear properties of the galaxy
density field or the reconstructuted dark matter density field.

The emergence of this web like structure is a great success of large
scale structure growth models based on gravitational
instability. Using numerical simulations allows the time description
of such structure together with the formation of the virialized dark
matter halos, which in turn should host observable galaxies. The
finding of gas filaments that feed galaxies is another theoretical
hint that strenghten the expected connection between galactic
properties ant their place in the cosmic web.

Using high resolution simulations it has been stablished that halos of
a given mass for earlier in denser regions; concentration, spin and
shape also depend on the local density. In the last decade new
algorithms have implemented cosmic web classifications that go beyond
the local density by defining a location to be a peak, sheet, filament
or void depending on the simmetry properties of the local dark matter
distribution.

There analytical frameworks, such as tidal torque theory, that
aim at describing how galaxies gain angular momentum and expect a
correlation of this quantity with the cosmic web. The observational
studies that try to measure such correlations actually try to use as a
proxy the observed shape of galaxies. In this respect it is useful to
have a theoretical baseline for the correlations of angular momentum
and shape with the cosmic web.

In the same spirit of describing the place of galaxies within the
cosmic web, there has been an revival of surveys that measure the
cosmic flow patters in the Local Universe. Assuming the correspondence
between the divergenec of the cosmic flow velocity field and the local
matter overdensity (valid in the linear regime) one could construct
accurate maps of the matter density from peculiar velocities. From
this perspective is interesting to look at the expected alignment of
the peculiar velocities.

There is large tradition of alignment measurements of shape and
angular momentum. The main result of these previous studies is that shape
alignment is a robust measurement regardless of the methods and
simulations. On the other hand, the results for the angular momentum
differ in the degree of the alignment. In this \documentname we review
most of these studyes and offer our own study with complementary
numerical techniques and simulations.  We also present for the first
time in the literature new results for the alignment of peculiar velocities
with large scale structure.

@@

\begin{figure*}
\includegraphics[width=0.8\textwidth]{Fig1.pdf}
\caption{Fraction of halos in a mass bin that show a preferential
  alignment with respect to an eigenvector. Each row presents on of
  the three properties studied in this paper: shape (major axis),
  angular momentum and  peculiar velocity. Left column presents the
  results against the Vweb   and the right the Tweb. Strong colors
  refer to $256^3$ grid resolutions and lighter colors to a $512^3$
  grid. The thick black horizontal line at $0.33$ corresponds to the
  expected fraction for a random vector field.
\label{fig:preferential}}
\end{figure*}

@@

\begin{figure}
\includegraphics[width=0.40\textwidth]{Fig5.pdf}
\caption{Median of the interweb alignment for the two grid
  resolutions as a function of the dark matter mass corresponding to
  the halos where the measurement was made. The error bars indicate
  the lower and upper quartiles (Figure in color in the web version).
\label{fig:interweb}} 
\end{figure}

@@

{\bf Possible conclusion. Prolateness follow concentration. In the Tweb
Angular momentum alignment is influenced by spin at higher masses. In
the Vweb angular momentum alginment is influenced by all factors. In
the peculiar velocity alginment only spin seemss to have an effect. }

\begin{figure*}
\includegraphics[width=0.90\textwidth,angle=90]{Fig6.pdf}
\caption{The alignment is always taken with respect to the third eigenvector.}
\end{figure*}
\section{Discussion}
\label{sec:discussion}
